---
output: github_document
---

# Predicting Flu Hospitalizations Paper

## Installation

Before installing this package, users will need to install the `glmgen` package, which is not on CRAN but is a dependency:

```{r, eval = F}
library(remotes)
remotes::install_github("statsmaths/glmgen", subdir = "R_pkg/glmgen")
remotes::install_github("jrgant/FluHospPrediction")
```

## Package directory (two levels deep):

```{r, echo = F}
fs::dir_tree(recurse = 2, regexp = "interim", invert = T)
```


## Code files

- All analytic code files are stored in /inst directory.
- Functions to handle calendar date management and curve simulation located in /R

## Data Sources

### FluSurv-NET

Chaves SS, Lynfield R, Lindegren ML, Bresee J, Finelli L. **The US Influenza Hospitalization Surveillance Network**. _Emerg Infect Dis._ 2015 Sep;21(9):1543–50. Available from: http://dx.doi.org/10.3201/eid2109.141912

**FluView: Influenza Hospitalization Surveillance Network**, Centers for Disease Control and Prevention. WEBSITE. (Emerging Infections Program data)


### WHO/NREVSS Viral Activity

National Center for Immunization and Respiratory Diseases (NCIRD). **U.S. Influenza Surveillance System: Purpose and Methods** [Internet]. Centers for Disease Control and Prevention. 2019 [cited 2020 Jan 9]. Available from: https://www.cdc.gov/flu/weekly/overview.htm


### ILINet

National Center for Immunization and Respiratory Diseases (NCIRD). **U.S. Influenza Surveillance System: Purpose and Methods** [Internet]. Centers for Disease Control and Prevention. 2019 [cited 2020 Jan 9]. Available from: https://www.cdc.gov/flu/weekly/overview.htm


### USCRN Temperature and Relative Humidity

Diamond HJ, Karl TR, Palecki MA, Baker CB, Bell JE, Leeper RD, et al. **U.S. Climate Reference Network after One Decade of Operations: Status and Assessment**. _Bull Am Meteorol Soc_ 2013;94:485–98. doi:10.1175/BAMS-D-12-00170.1.


### Historical Thanksgiving Dates

Wikipedia contributors. **Thanksgiving (United States)**. In: _Wikipedia, The Free Encyclopedia_ [Internet]. 30 Dec 2019 [cited 10 Jan 2020]. Available: https://en.wikipedia.org/w/index.php?title=Thanksgiving_(United_States)&oldid=933219107
