---
output: github_document
---

# Predicting Flu Hospitalizations Paper

## Installation

Before installing this package, users will need to install the `glmgen` and `sl3` packages, which are not on CRAN but are dependencies:

```{r, eval = F}
library(remotes)
remotes::install_github("statsmaths/glmgen", subdir = "R_pkg/glmgen")
remotes::install_github("tlverse/sl3")
remotes::install_github("jrgant/FluHospPrediction")
```

Please post an issue if you have trouble installing.


## Package directory (two levels deep):

```{r, echo = F}
fs::dir_tree(recurse = 2, regexp = "interim|results", invert = T)
```


## Code files

- All analytic code files are stored in /inst directory.
- Functions to handle calendar date management and curve simulation located in /R

## Data Sources

### FluSurv-NET

Chaves SS, Lynfield R, Lindegren ML, Bresee J, Finelli L. **The US Influenza Hospitalization Surveillance Network**. _Emerg Infect Dis._ 2015 Sep;21(9):1543–50. Available from: http://dx.doi.org/10.3201/eid2109.141912

**FluView: Influenza Hospitalization Surveillance Network**, Centers for Disease Control and Prevention. WEBSITE. (Emerging Infections Program data)
